Continuous Reinforcement Learning

### 2018-09-21 magnitudes.tex
Changes.
Introducing A function.

The state value function and action-state value function under policy
$\pi$ are defined as
\begin{align}
	Q^\pi(s, a) &= \BE_\pi\left[
		\sum\limits_{k=0}^\infty
	{\gamma}^{k \deltat} r(s_k, a_k)
		\mid
		s_0 = s, a_0 = a
	\right]\\
	&= \BE_\pi\left[
		\sum\limits_{k=0}^\infty
		{\gamma}^{k \deltat} \tilde{r}(s_k, a_k) \deltat
		\mid
		s_0 = s, a_0 = a
	\right]\\
	V^\pi(s) &= \BE_\pi\left[Q^\pi(s_t, a_t) | s_t=s\right].
\end{align}
Both functions satisfy recurrent \emph{Bellman equations}, e.g.
\begin{align}
	Q^\pi(s, a) &= \BE_\pi\left[
		\tilde{r}(s_t, a_t) \deltat + \gamma^\deltat V^\pi(s_{t + \deltat})
		\mid
	s_t = s, a_t = a\right]
\end{align}
as well as $K$ steps \emph{Bellman equations}, for any $K$, and notably for
$K = \frac{1}{\deltat}$, e.g.
\begin{align}
	V^\pi(s) = \BE_\pi\left[
		\sum\limits_{k=0}^{\frac{1}{\deltat} - 1} \gamma^{k\deltat} \tilde{r}(s_k, a_k) \deltat + \gamma V^\pi(s_{t + 1})
		\mid
	s_0 = s\right].
\end{align}
With a fixed policy $\pi$, the trajectory of states in rollouts is a
\emph{Markov Chain}. Under mild conditions, there exists a stationnary
distribution $\rho$ for this markov chain. With those notations, the goal to
maximize, the average return, is defined as
$\BE_{s\sim\rho}\left[V^{\pi}(s)\right]$.
Optimal $V$ and $Q$ functions, i.e. $V^*=V^{\pi^*}$ and $Q^*=Q^{\pi^*}$ with
$\pi^*$ the policy that maximizes the average return, also satisfy Bellman equations
equations, i.e.
\begin{align}
	Q^*(s, a) &= \BE\left[
		\tilde{r}(s, a) \right] \deltat + \gamma^\deltat \max\limits_{a'} \BE\left[Q^*(s_{t + \deltat}, a')
		\mid
	s_t = s, a_t = a\right].
	\label{eq:optq}
\end{align}
Popular approaches to optimizing
the average return mainly fall in two categories:
\begin{itemize}
	\item $Q$-learning approaches, which use~\eqref{eq:optq} as a fixed point equation
		to derive iterative approximations of the optimal $Q$ function.
	\item Policy gradient approaches, which search through a parameterized-sub family of
		all policies in order to optimize the average return through gradient descent. 
		It is possible to estimate the gradient of the average return
		using the policy gradient theorem, which yields
		\begin{equation}
			\nabla_\theta\BE_{s\sim\rho_\theta}\left[V^{\pi_\theta}(s)\right]
			=
			\BE_{s\sim\rho_\theta, a\sim\pi(\cdot\mid s)}\left[
				Q^{\pi_\theta}(s, a)
				\nabla_\theta \log \pi_\theta(a \mid s)
			\right].
		\end{equation}
\end{itemize}
Both approaches require estimating $Q^{\pi_\theta}$, and we will show that these
estimations need to be precise up to $\bigO{\deltat}$.

@@

Approximating with this degree of precision becomes harder as $\bigO{\deltat}$
decreases. Typically, if the Q-function is learnt only from \emph{Monte Carlo} samples
of the return, the observed variance of the quantity to learn is of order
$\frac{1}{1 - \gamma}$. An optimist estimate of the number of
samples required to learn this quantity up to order $\bigO{\deltat}$ is 
$\bigO{\frac{1}{(1 - \gamma) \deltat^2}}$. If a sampling rate of $\deltat$ is
assumed, the time it takes to learn $Q$ up to the correct degree of precission,
increases linearily with $\frac{1}{\deltat}$. Indeed, in a single real time step, with a
sampling rate of $\deltat$, $\frac{1}{\deltat}$ samples are aquired. Consequently,
$\frac{1}{(1 - \gamma) \deltat}$ timesteps are required to obtain an estimate up
to the desired precision. As $\deltat \to 0$, it takes an infinite number of samples
to obtain a viable estimate of $Q$, rendering $Q$-learning methods theoretically
unsound.
% \NDC{This is not convincing at the moment: as $\deltat$ decreases,
% 	the number of observations made increases, and this may help get a better
% 	estimate $\tilde{Q}^\pi(s, a)$. My guess is that the number of observations
% 	required to obtain an approximation of order $\varepsilon$ is superlinear in
% 	$\frac{1}{\varepsilon}$ and thus the previous argument holds. We either need
% a simple theoretical setup where this works or an experimental one.}
% }}}

\paragraph{\emph{Monte Carlo} policy gradient in continuous time}~\\
% {{{
% }}}

@@

\paragraph{Questions left unanswered with advantage learning.}
% {{{
As mentionned, advantage learning requires learning the gradient of $V$
with respect to the states, or, in the discrete case, the difference of
$V$'s between two neighboring rescaled by $\frac{1}{\deltat}$. Now, as
mentionned in~\ref{sec:QContinuous}, $V$ can only be learned up to $\bigO{1}$
in reasonable time. Consequently, $V(s) - V(s') = \bigO{1}$ in the general case,
and $\frac{V(s) - V(s')}{\deltat} = \bigO{\frac{1}{\deltat}}$. This means that
the learning procedure of $A$ can endure a bias of order $\bigO{\frac{1}{\deltat}}$
which goes to $+\infty$ as $\deltat \to 0$.

>>>>

The state value function and action-state value function under policy
$\pi$ are defined as
\begin{align}
	Q^\pi(s, a) &= \BE_\pi\left[
		\sum\limits_{k=0}^\infty
	{\gamma}^{k \deltat} r(s_k, a_k)
		\mid
		s_0 = s, a_0 = a
	\right]\\
	&= \BE_\pi\left[
		\sum\limits_{k=0}^\infty
		{\gamma}^{k \deltat} \tilde{r}(s_k, a_k) \deltat
		\mid
		s_0 = s, a_0 = a
	\right]\\
	V^\pi(s) &= \BE_\pi\left[Q^\pi(s_t, a_t) | s_t=s\right].
\end{align}
Both functions satisfy recurrent \emph{Bellman equations}, e.g.
\begin{align}
	Q^\pi(s, a) &= \BE_\pi\left[
		\tilde{r}(s_t, a_t) \deltat + \gamma^\deltat V^\pi(s_{t + \deltat})
		\mid
	s_t = s, a_t = a\right]
\end{align}
as well as $K$ steps \emph{Bellman equations}, for any $K$, and notably for
$K = \frac{1}{\deltat}$, e.g.
\begin{align}
	V^\pi(s) = \BE_\pi\left[
		\sum\limits_{k=0}^{\frac{1}{\deltat} - 1} \gamma^{k\deltat} \tilde{r}(s_k, a_k) \deltat + \gamma V^\pi(s_{t + 1})
		\mid
	s_0 = s\right].
\end{align}
With a fixed policy $\pi$, the trajectory of states in rollouts is a
\emph{Markov Chain}. Under mild conditions, there exists a stationnary
distribution $\rho$ for this markov chain. With those notations, the goal to
maximize, the average return, is defined as
$\BE_{s\sim\rho}\left[V^{\pi}(s)\right]$.
Optimal $V$ and $Q$ functions, i.e. $V^*=V^{\pi^*}$ and $Q^*=Q^{\pi^*}$ with
$\pi^*$ the policy that maximizes the average return, also satisfy Bellman equations
equations, i.e.
\begin{align}
	Q^*(s, a) &= \BE\left[
		\tilde{r}(s, a) \right] \deltat + \gamma^\deltat \max\limits_{a'} \BE\left[Q^*(s_{t + \deltat}, a')
		\mid
	s_t = s, a_t = a\right].
	\label{eq:optq}
\end{align}
Popular approaches to optimizing
the average return mainly fall in two categories:
\begin{itemize}
	\item $Q$-learning approaches, which use~\eqref{eq:optq} as a fixed point equation
		to derive iterative approximations of the optimal $Q$ function.
	\item Policy gradient approaches, which search through a parameterized-sub family of
		all policies in order to optimize the average return through gradient descent. 
		It is possible to estimate the gradient of the average return
		using the policy gradient theorem, which yields
		\begin{equation}
			\label{eq:policygradient}
			\nabla_\theta\BE_{s\sim\rho_\theta}\left[V^{\pi_\theta}(s)\right]
			=
			\BE_{s\sim\rho_\theta, a\sim\pi(\cdot\mid s)}\left[
				Q^{\pi_\theta}(s, a)
				\nabla_\theta \log \pi_\theta(a \mid s)
			\right].
		\end{equation}
\end{itemize}
Both approaches require estimating $Q^{\pi_\theta}$, and we will show that these
estimations need to be precise up to $\bigO{\deltat}$.

@@

Approximating with this degree of precision becomes harder as $\bigO{\deltat}$
decreases. Typically, if the Q-function is learnt only from \emph{Monte Carlo} samples
of the return, the observed variance of the quantity to learn is of order
$\frac{1}{1 - \gamma}$. An optimist estimate of the number of
samples required to learn this quantity up to order $\bigO{\deltat}$ is 
$\bigO{\frac{1}{(1 - \gamma) \deltat^2}}$. If a sampling rate of $\deltat$ is
assumed, the time it takes to learn $Q$ up to the correct degree of precission,
increases linearily with $\frac{1}{\deltat}$. Indeed, in a single real time step, with a
sampling rate of $\deltat$, $\frac{1}{\deltat}$ samples are aquired. Consequently,
$\frac{1}{(1 - \gamma) \deltat}$ timesteps are required to obtain an estimate up
to the desired precision. As $\deltat \to 0$, it takes an infinite number of samples
to obtain a viable estimate of $Q$, rendering $Q$-learning methods theoretically
unsound.

In what follows, we will refer to the advantage function $A^\pi(s, a)$ as
$\frac{Q^\pi(s,a) - V^\pi(s)}{\deltat}$. Since $Q^\pi$ and $V^\pi$ only differ
by a factor of scale $\bigO{\deltat}$, this quantity remains finite as $\deltat \to 0$.
% \NDC{This is not convincing at the moment: as $\deltat$ decreases,
% 	the number of observations made increases, and this may help get a better
% 	estimate $\tilde{Q}^\pi(s, a)$. My guess is that the number of observations
% 	required to obtain an approximation of order $\varepsilon$ is superlinear in
% 	$\frac{1}{\varepsilon}$ and thus the previous argument holds. We either need
% a simple theoretical setup where this works or an experimental one.}
% }}}

\paragraph{Policy gradient methods in continuous time} Policy gradient methods aim at
maximizing the average return using gradient ascent on estimations of~\eqref{eq:policygradient}.
One can observe that subtracting an action independent baseline to $Q^{\pi_\theta}$ in~\eqref{eq:policygradient} still yields an unbiased estimte of the gradient. Notably,
\begin{equation}
\end{equation}

% {{{
% }}}

@@

\paragraph{Questions left unanswered with advantage learning.}
% {{{
As mentionned, advantage learning requires learning the gradient of $V$
with respect to the states, or, in the discrete case, the difference of
$V$'s between two neighboring rescaled by $\frac{1}{\deltat}$. Now, as
mentionned in~\ref{sec:QContinuous}, $V$ can only be learnt up to $\bigO{1}$
in reasonable time. Consequently, $V(s) - V(s') = \bigO{1}$ in the general case,
and $\frac{V(s) - V(s')}{\deltat} = \bigO{\frac{1}{\deltat}}$. This means that
the learning procedure of $A$ can endure a bias of order $\bigO{\frac{1}{\deltat}}$
which goes to $+\infty$ as $\deltat \to 0$.
